---
title: "Evaluating the Classical Linear Model Assumptions"
subtitle: "Q1.5 Normally Distributed Errors"
output: "pdf_document"
---
```{r install packages, message=FALSE, echo=FALSE, include=FALSE}
if(!require(pacman)){install.packages("pacman")}
library(pacman)
p_load(tidyverse, patchwork, lmtest, sandwich, stargazer)


```

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message=FALSE)
library(tidyverse)
library(patchwork)
library(lmtest)
library(sandwich)
library(stargazer)
theme_set(theme_minimal())

```

```{r load data, include=FALSE}

videos <- read_delim("./videos.txt")

videos <- videos %>% na.omit()

```

First the linear regression model was created.

```{r model creation, echo=TRUE}

model_12 <- lm(log(views) ~ rate + length, data= videos)

```

```{r}

videos <- videos %>%
  mutate(
    model_12_preds = predict(model_12),
    model_12_resids = resid(model_12)
  )

```

To assess if the errors are normally distributed a histogram and qq plot were created. The histogram of the residuals reveals a relatively normal distribution. Examination of the qq plot further supports a normal distribution with only slight deviation from the theoretical normal line indicating some thin tails.

```{r}

errors_dist <- videos %>%
  ggplot(aes(x= model_12_resids))+
  geom_histogram()


errors_qq <- videos %>%
  ggplot(aes(sample= model_12_resids))+
  stat_qq() + stat_qq_line()

errors_dist / errors_qq
```

Therefore, we conclude that the assumption for Normally Distributed Errors is met.