---
title: "Evaluating the Classical Linear Model Assumptions"
subtitle: "Q1.4 Constant Error Variance"
output: "pdf_document"
---
```{r install packages, message=FALSE, echo=FALSE, include=FALSE}
if(!require(pacman)){install.packages("pacman")}
library(pacman)
p_load(tidyverse, patchwork, lmtest, sandwich, stargazer)


```

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message=FALSE)
library(tidyverse)
library(patchwork)
library(lmtest)
library(sandwich)
library(stargazer)
theme_set(theme_minimal())

```

```{r load data, include=FALSE}

videos <- read_delim("./videos.txt")

videos <- videos %>% na.omit()

```

First the linear regression model was created.

```{r model creation, echo=TRUE}

model_12 <- lm(log(views) ~ rate + length, data= videos)

```

```{r}

videos <- videos %>%
  mutate(
    model_12_preds = predict(model_12),
    model_12_resids = resid(model_12)
  )

preds_resids <- videos %>%
  ggplot(aes(model_12_preds,model_12_resids)) +
  geom_point()+
  stat_smooth()+
  labs(
    x= "Model Predictions",
    y= "Model Residuals"
  )

```

To assess whether the distribution of the errors is homoskedastic, we can perform an ocular test on the model residuals versus predicted values plot. Homoskedastic errors would reveal a band of even thickness from left to right. Examining the plot reveals a slightly uneven band in the lower half of the predicted values though it is not severe. However in the upper half of the predicted values there is a sharp decrease in the variance of residuals.

```{r analysis of errors, fig.height=4.0, fig.align='center'}

preds_resids

```
Additionally, we can conduct a Breusch-Pagan test. A p-value less than 0.05 would indicate heteroskedasticity. Analysis of the test below reveals a statistically significant result supporting heteroskedasticity.
```{r}
lmtest::bptest(model_12)
```

 Therefore, we conclude the assumption for Constant Error Variance / Homoskedasticity of Errors is not met.

