---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message=FALSE)

library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(message=FALSE)
theme_set(theme_minimal())
```

```{r}
df = read.table(file = 'data/videos.txt', sep = '\t', header = TRUE)
# Remove NaN values
df <- df %>% drop_na(rate)
df <- df %>% drop_na(length)
df <- df %>% drop_na(views)
```

Building the model: lm(log(views) ~ length + rate, data = df)
```{r}
model_1 <- lm(log(views) ~ length + rate, data = df)

df <- df %>%  
  mutate(model_1_residuals = resid(model_1),
         model_1_predictions = predict(model_1))
```

Plotting the model predictions vs residuals to perform an ocular test to verify homoscedasticity.
```{r}
plot_ocular <- df %>% 
  ggplot(aes(x = model_1_predictions, y = model_1_residuals)) + 
  geom_point() + stat_smooth(se = TRUE)

plot_ocular
```

From the above plot we can see that the model tends to under-predict for larger predictions, leading to negative residuals for upper range of the plot on the x.

Conducting the Breusch-Pagan test to look for evidence of heteroscedasticity.
```{r}
lmtest::bptest(model_1)
```
We get a very small p-value, so we have strong evidence to reject the null hypothesis and conclude that there is some heteroscedasticity in the way residuals are distributed.


