---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message=FALSE)

library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(message=FALSE)
theme_set(theme_minimal())
```

```{r}
df = read.table(file = 'data/videos.txt', sep = '\t', header = TRUE)
# Remove NaN values
df <- df %>% drop_na(rate)
df <- df %>% drop_na(length)
df <- df %>% drop_na(views)
```

Building the model: lm(log(views) ~ length + rate, data = df)

```{r}
model_1 <- lm(log(views) ~ length + rate, data = df)

df <- df %>%  
  mutate(model_1_residuals = resid(model_1),
         model_1_predictions = predict(model_1))
```

Plotting a histogram to view the distribution of the residuals.
```{r}
# Draw the histogram
hist(df$model_1_residuals)
```
From a glance it looks like the distribution is roughly normal and distributed quite evenly about 0.

Plotting the qq-plot for the residuals.
```{r}
# Draw the Q-Q plot
qqnorm(df$model_1_residuals)
qqline(df$model_1_residuals, col = "blue", lwd = 2)
```
Similar to our histogram, from the above plot it looks like the values are normally distributed with only slight deviations near the ends. Normality of errors is met.
