---
title: "R Notebook"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message=FALSE)

library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(message=FALSE)
theme_set(theme_minimal())
```

```{r}
df = read.table(file = 'data/videos.txt', sep = '\t', header = TRUE)
# Remove NaN values
df <- df %>% drop_na(rate)
df <- df %>% drop_na(length)
df <- df %>% drop_na(views)
```

Building the model: lm(log(views) ~ length + rate, data = df)
```{r}
model_1 <- lm(log(views) ~ length + rate, data = df)
```

Correlation of length and rate columns:
```{r}
cor(df$length, df$rate)
```

Getting model summary:
```{r}
summary(model_1)
```

Neither feature was dropped, so there must be no perfect collinearity. Also, the p-value assigned to both features is very low, so they're both important to our regression. Furthermore, the correlation between length and rate is only 0.16, so one is not a linear combination of the other.



